---
title: "math_motion_project"
output: html_document
date: '2023-11-16'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#library(renv)
#renv::init()

#library(ggplot2)

library(tidyverse)
library(plotly)

#renv::snapshot()

```


```{r}
## read and understand data ##

math.df <- read.csv('coordinates_Dami.csv')
```

```{r}
#checking the head and the tail of my dataset

#head(math.df)
#tail(math.df)

```


```{r}
#looking at the column's info
#str(time.df)

```


```{r}
#time (column X) is in hertz. 
#Adding a column for time in seconds (1 second = 30 hertz) 

math.df <- math.df %>%
    mutate(t_sec = X/30) 

```

# Shoulder motion

## Machine's confidnece in capturing the right motion of shoulder

```{r}
#V stands for visibility of the body parts (the red line) and how confident the machine is in capturing the right motion.

TS_right_shoulder <- math.df %>%
    ggplot(aes(x= t_sec, y= RIGHT_SHOULDER_Y)) +
    geom_point() +
    geom_path() +
    geom_path(aes(y=RIGHT_SHOULDER_V), color="red") +
    geom_path(aes(y=RIGHT_SHOULDER_X), color="blue") +
    scale_x_continuous("time (sec.)", expand=c(0,0)) +
    theme_classic()

#ggsave("TS_right_shoulder_plot.png", TS_right_shoulder)

TS_right_shoulder

```

## the plot above shows that the machine algorithm tracks participant's shoulder confidently (V = 1 throughout the session).

```{r}
#the insight (aha!) moment (observable by words such as "oh, OK!") happened at the second 203 based on the video recording

```


```{r}
#plotting X and Y of right shoulder motion

right_shoulder_plot <- ggplot(math.df, aes(x = RIGHT_SHOULDER_X, y =  RIGHT_SHOULDER_Y)) +
    geom_point() +
    geom_path() +
    theme_classic()
right_shoulder_plot

```
## The above plot shows that X is the mian (or at least more important) factor for distance

```{r}
#plotting shoulder's x axes over time

right_shoulder_x <- ggplot(math.df, aes(x = X, y =  RIGHT_SHOULDER_X)) +
  geom_point() +
  geom_path() +
  geom_vline(xintercept = 203*30, linetype = "dashed", color = "red")+  
  
  xlab("time")+
  theme_classic()

right_shoulder_x

#ggsave("right_shoulder_x.png", right_shoulder_x )

```



```{r}
#min(time.df.sec$RIGHT_SHOULDER_X)
#max(time.df.sec$RIGHT_SHOULDER_X)

# Create labels for categorical distance
labels <- c("close", "medium", "far")

a <- 0.16 # less than a = close distance
b <- 0.26 # greater than b = far distance


df.shoulder.c <- math.df %>% select(X, t_sec, RIGHT_SHOULDER_X, RIGHT_SHOULDER_Y, RIGHT_SHOULDER_Z, RIGHT_SHOULDER_V)
# Use cut to create the categorical variable
df.shoulder.c$RIGHT_SHOULDER_X <- cut(df.shoulder.c$RIGHT_SHOULDER_X, breaks = c(-Inf, a, b, Inf), labels = labels, right = FALSE)

# Print the result
#df.shoulder.c

```

```{r}
# Preprocess the data to create a new variable for time slots
t.lables_x = c("t1", "t2", "t3", "t4", "t5", "t6")
df.shoulder.c2 <- df.shoulder.c %>%
  mutate(TimeSlot_x = cut(X, breaks = 6, labels = t.lables_x)) #

# Create the plot
ggplot(df.shoulder.c2, aes(x = TimeSlot_x, fill = RIGHT_SHOULDER_X)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  labs(title = "Distribution of RIGHT SHOULDER X over Time Slots",
       x = "Time Slot",
       y = "Proportion") +
  theme_minimal()

```




```{r}
#plotting shoulder's y axes over time

right_shoulder_y <- ggplot(math.df, aes(x = X, y =  RIGHT_SHOULDER_Y)) +
  geom_point() +
  geom_path() +
  geom_vline(xintercept = 203*30, linetype = "dashed", color = "red") +
  xlab("time")+
  
  theme_classic()
right_shoulder_y


```



```{r}
#min(time.df.sec$RIGHT_SHOULDER_X)
#max(time.df.sec$RIGHT_SHOULDER_X)

# Create labels for categorical distance
labels_y <- c("close", "medium", "far")

aa <- 0.56 # less than aa = close distance
bb <- 0.58 # greater than bb = far distance

#df.shoulder.c <- math.df %>% select(X, t_sec, RIGHT_SHOULDER_X, RIGHT_SHOULDER_Y, RIGHT_SHOULDER_Z, RIGHT_SHOULDER_V)

# Use cut to create the categorical variable
df.shoulder.c$RIGHT_SHOULDER_Y <- cut(df.shoulder.c$RIGHT_SHOULDER_Y, breaks = c(-Inf, aa, bb, Inf), labels = labels_y, right = FALSE)

# Print the result
df.shoulder.c

```

```{r}
# Preprocess the data to create a new variable for time slots
t.lables_y = c("t1", "t2", "t3", "t4", "t5", "t6")
 df.shoulder.c2 <- df.shoulder.c %>%
   mutate(TimeSlot_y = cut(X, breaks = 6, labels = t.lables_y)) #

# Create the plot
ggplot(df.shoulder.c2, aes(x = TimeSlot_y, fill = RIGHT_SHOULDER_Y)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  labs(title = "Distribution of RIGHT SHOULDER Y over Time Slots",
       x = "Time Slot",
       y = "Proportion") +
  theme_minimal()

#ggsave("shoulder_Y_distancePlot.png")


```





# right wrist motion

## Machine's confidnece in capturing the right motion of wrist

```{r}
#V stands for visibility of the body parts (the red line) and how confident the machine is in capturing the right motion.

TS_right_wrist <- math.df %>%
    ggplot(aes(x= t_sec, y= RIGHT_WRIST_Y)) +
    geom_point() +
    geom_path() +
    geom_path(aes(y=RIGHT_WRIST_V), color="red") +
    geom_path(aes(y=RIGHT_WRIST_X), color="blue") +
    scale_x_continuous("time (sec.)", expand=c(0,0)) +
    theme_classic()

TS_right_wrist

#ggsave("TS_right_wrist.png", TS_right_wrist)

```

## the plot above shows that the machine algorithm doest not track participant's shoulder very confidently.


```{r}
#plotting X and Y of right wirst motion

right_wrist_plot <- ggplot(math.df, aes(x = RIGHT_WRIST_X, y =  RIGHT_WRIST_Y)) +
    geom_point() +
    geom_path() +
    theme_classic()
right_wrist_plot

```
## The above plot shows that both X and Y important factor for distance.

```{r}
#plotting wrist's x axes over time

right_wrist_x <- ggplot(math.df, aes(x = X, y =  RIGHT_WRIST_X)) +
  geom_point() +
  geom_path() +
  geom_vline(xintercept = 203*30, linetype = "dashed", color = "red")+  
  
  xlab("time")+
  theme_classic()

right_wrist_x


```


```{r}
#min(math.df$RIGHT_WRIST_X)
#max(math.df$RIGHT_WRIST_X)

# Create labels for categorical distance
#labels <- c("close", "medium", "far")

r1 <- 0.17 # less than r1 = close distance
r2 <- 0.30 # greater than r2 = far distance

df.wrist.c <- math.df %>% select(X, t_sec, RIGHT_WRIST_X, RIGHT_WRIST_Y, RIGHT_WRIST_Z, RIGHT_WRIST_V)
# Use cut to create the categorical variable
df.wrist.c$RIGHT_WRIST_X <- cut(df.wrist.c$RIGHT_WRIST_X, breaks = c(-Inf, r1, r2, Inf), labels = labels, right = FALSE)

# Print the result
df.wrist.c

```

```{r}
# Preprocess the data to create a new variable for time slots
#t.lables_x = c("t1", "t2", "t3", "t4", "t5", "t6")
df.wrist.c2 <- df.wrist.c %>%
  mutate(TimeSlot_x = cut(X, breaks = 6, labels = t.lables_x)) #

# Create the plot
ggplot(df.wrist.c2, aes(x = TimeSlot_x, fill = RIGHT_WRIST_X)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  labs(title = "Distribution of RIGHT SHOULDER X over Time Slots",
       x = "Time Slot",
       y = "Proportion") +
  theme_minimal()

```




```{r}
#plotting wrist's y axes over time

right_wrist_y <- ggplot(math.df, aes(x = X, y =  RIGHT_WRIST_Y)) +
  geom_point() +
  geom_path() +
  geom_vline(xintercept = 203*30, linetype = "dashed", color = "red") +
  xlab("time")+
  
  theme_classic()

right_wrist_y


```



```{r}
#min(time.df.sec$RIGHT_WRIST_Y)
#max(time.df.sec$RIGHT_WRIST_Y)

# Create labels for categorical distance
#labels_y <- c("close", "medium", "far")

ry1 <- 0.5 # less than ry1 = close distance
ry2 <- 0.7 # greater than ry2 = far distance

#df.wrist.c <- math.df %>% select(X, t_sec, RIGHT_WRIST_X, RIGHT_WRIST_Y, RIGHT_WRIST_Z, RIGHT_WRIST_V)

# Use cut to create the categorical variable
df.wrist.c$RIGHT_WRIST_Y <- cut(df.wrist.c$RIGHT_WRIST_Y, breaks = c(-Inf, ry1, ry2, Inf), labels = labels_y, right = FALSE)

# Print the result
df.wrist.c

```

```{r}
# Preprocess the data to create a new variable for time slots
# t.lables_y = c("t1", "t2", "t3", "t4", "t5", "t6")
 df.wrist.c2 <- df.wrist.c %>%
   mutate(TimeSlot_y = cut(X, breaks = 6)) #, labels = t.lables_y

# Create the plot
ggplot(df.wrist.c2, aes(x = TimeSlot_y, fill = RIGHT_WRIST_Y)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  labs(title = "Distribution of RIGHT WRIST Y over Time Slots",
       x = "Time Slot",
       y = "Proportion") +
  theme_minimal()


```

#Principle Component Analysis for right wrist

```{r}
#preparing data for Priciple Component Analysis
wrist.df = math.df %>% select(RIGHT_WRIST_X, RIGHT_WRIST_Y, RIGHT_WRIST_Z)

```

```{r}
pca_wrist = prcomp(wrist.df, center = T, scale. = T)

summary(pca_wrist)
```

```{r}
#Extract principal components
pc_scores <- pca_wrist$x  # Scores (coordinates of the data points in the new space)
pc_loadings <- pca_wrist$rotation  # Loadings (coefficients of the original variables in the principal components)
```

```{r}
#Visualizing the proportion of variance explained by each principal component
plot(cumsum(pca_wrist$sdev^2) / sum(pca_wrist$sdev^2), xlab = "Number of Principal Components", ylab = "Cumulative Proportion of Variance Explained", type = "b")

```


```{r}
#based on the plot above by two components 90% of the variance can be explained 
#Reduce dimensions
num_components_to_keep <- 2  # Choosing the number of components we want to keep
reduced_data <- pc_scores[, 1:num_components_to_keep]

# The 'reduced_data' dataframe now contains the data with reduced dimensions

```


```{r}
# Create a dataframe for the reduced data
reduced_data_df <- as.data.frame(reduced_data)

# Scatter plot of the reduced data
plot(reduced_data_df$PC1, reduced_data_df$PC2, 
     xlab = "Principal Component 1",
     ylab = "Principal Component 2",
     main = "Scatter Plot of Reduced Data")


```


```{r}
#plotting a 3D plot of the data

plot_3D_wrist <- plot_ly(
  data = wrist.df,
  x = ~RIGHT_WRIST_X,
  y = ~RIGHT_WRIST_Y,
  z = ~RIGHT_WRIST_Z,
  #text = ~paste("Time: ", time),
  type = "scatter3d",
  mode = "markers"
)

# Customize the layout
layout <- list(
  scene = list(
    xaxis = list(title = 'X-axis'),
    yaxis = list(title = 'Y-axis'),
    zaxis = list(title = 'Z-axis')
  )
)

# Combine plot and layout
plot_3D_wrist <- plot_3D_wrist %>% layout(layout)

# Display the interactive plot
plot_3D_wrist
```















```{r}
#plotting shoulder's z axes over time

right_shoulder_z <- ggplot(math.df, aes(x = X, y =  RIGHT_SHOULDER_Z)) +
  geom_point() +
  geom_path() +
  geom_vline(xintercept = 203*30, linetype = "dashed", color = "red") +
  xlab("time")+
  
  theme_classic()
right_shoulder_z


```

#Principle Component Analysis 

```{r}
#preparing data for Priciple Component Analysis
shoulder.df = math.df %>% select(RIGHT_SHOULDER_X, RIGHT_SHOULDER_Y, RIGHT_SHOULDER_Z)
#shoulder.df
```

```{r}
pca_shoulder = prcomp(shoulder.df, center = T, scale. = T)

summary(pca_shoulder)
```

```{r}
#Extract principal components
pc_scores <- pca_shoulder$x  # Scores (coordinates of the data points in the new space)
pc_loadings <- pca_shoulder$rotation  # Loadings (coefficients of the original variables in the principal components)
```

```{r}
#Visualizing the proportion of variance explained by each principal component
plot(cumsum(pca_shoulder$sdev^2) / sum(pca_shoulder$sdev^2), xlab = "Number of Principal Components", ylab = "Cumulative Proportion of Variance Explained", type = "b")

```


```{r}
#based on the plot above by two components 90% of the variance can be explained 
#Reduce dimensions
num_components_to_keep <- 2  # Choosing the number of components we want to keep
reduced_data <- pc_scores[, 1:num_components_to_keep]

# The 'reduced_data' dataframe now contains the data with reduced dimensions

```


```{r}
# Create a dataframe for the reduced data
reduced_data_df <- as.data.frame(reduced_data)

# Scatter plot of the reduced data
plot(reduced_data_df$PC1, reduced_data_df$PC2, 
     xlab = "Principal Component 1",
     ylab = "Principal Component 2",
     main = "Scatter Plot of Reduced Data")


```


```{r}
#plotting a 3D plot of the data

plot_3D_shoulder <- plot_ly(
  data = shoulder.df,
  x = ~RIGHT_SHOULDER_X,
  y = ~RIGHT_SHOULDER_Y,
  z = ~RIGHT_SHOULDER_Z,
  #text = ~paste("Time: ", time),
  type = "scatter3d",
  mode = "markers"
)

# Customize the layout
layout <- list(
  scene = list(
    xaxis = list(title = 'X-axis'),
    yaxis = list(title = 'Y-axis'),
    zaxis = list(title = 'Z-axis')
  )
)

# Combine plot and layout
plot_3D_shoulder <- plot_3D_shoulder %>% layout(layout)

# Display the interactive plot
plot_3D_shoulder
```


